Updating the Feature Function Framework \titlelinebreak{} in the Moses Decoder

### 2013-07-18 mt-marathon.2013/PBML_article.tex
merge

\title{A Comparison of Synchronous-CFG \titlelinebreak{} Decoder Implementations}

@@

The open-source community has also changed from the dominance of the phrase-based model and virtual Moses monopoly, to a situation where syntactic models are leading the research agenda, and other projects such as cdec and Joshua have equal prominence.  We take this opportunity to compare the implementations of these other decoders with Moses. We find that there are many similarities between the decoders, but also some surprising differences in implementation of even very fundamental features in each decoder.

@@

Additionally, some feature functions do not override any of the stages. These feature is so intertwined with the search mechanism that their scores are set by search mechanism themselves. The feature function exist as an identifier to locate the score. Features functions such as the translation models, unknown word penalty,  generation models fall into this category. The input feature, which scores lattices and confusion network input is also hardcoded into the decoder.

@@

Joshua's decoding algorithm is an implementation of CKY+ \cite{chappelier1998generalized}. It
supports decoding with arbitrary synchronous CFG productions, imposing no limits on the rank of
rules or on the intermingling of terminals and nonterminals.

Joshua makes only a single CKY pass over the source language sentence; it does not build a
``stateless'' hypergraph (or --LM forest, in the language of \citet{chiang2007hierarchical}). To
enable this, Joshua interleaves the extension of partially-constructed rules with the application of
complete rules while building the hypergraph, depicted here with pseudocode:

@@

The efficiency gained from this interleaving allows Joshua to decode with SCFGs with rich
nonterminal sets \cite{galley2004whats,zollmann2006syntax}, although the time and memory
requirements increase dramatically. Incorporating feature scores into the dot chart for more
efficient decoding with SCFGs with rich nonterminal sets is a direction of interest.

\subsection{cdec}
In contrast to Moses and Joshua which use a one-pass search strategy, cdec uses a multi-pass search strategy. That is, for each input sentence, all feature functions are evaluated when a hypothesis is created in Moses.

In cdec, each pass creates a hypergraph that is used as the input for the next pass. cdec evaluates stateless feature functions in the first pass, followed by stateful feature functions in the second. Neither decoder partially evaluate stateful feature functions, as done in Moses.

@@

\subsection{Moses}
Moses implements supports for both dense and sparse features. Each feature must be identifiable as one or the other.
The intuition behind this is that most systems will still contain dense features. Therefore, it would be more efficient to use existing dedicated data-structures to efficiently handle dense vectors.

A sparse feature map was added to handle sparse scores.

@@

The results from the three decoders are comparable, given the same formalism and model data, and adjusting for scaling differences and constants.

@@

Figure~\ref{fig:glue} shows the improvement in cdec by changing the glue rule formulation from (2) to (1). The glue grammar for each decoder had varied since their inceptions. The convergence on using formulation (1) is the direct result of the comparison between the three decoders.

@@

\subsection{Head-to-head}

Figure~\ref{fig:model} shows the comparison of model scores and BLEU scores for a hierarchical model. Moses and cdec are roughly comparable in terms of speed. Also, Moses and cdec converges to roughly the same model scores, while Joshua makes more search errors, leading to lower BLEU scores.

@@

Statistical machine translation is a dynamic field which is constantly evolving. A major challenge for the Moses project, in common with many long-term software projects, is to maintain and update the the underlying framework and infrastructure to enable this evolution to continue. The paper documents the major changes that were made to Moses to enable this to happen and compare it two other decoders which implements the same decoding formalism.

Even though each decoder was created independently, their output are comparable. We see that by doing an indepth analysis of the working of each decoder and comparing their model scores, we are able to transfer the best optmization techniques from each decoder to the others.

>>>>

\title{Updating the Feature Function Framework \titlelinebreak{} in the Moses Decoder}

@@

We compare the implementation of the Moses SCFG decoder to that of cdec and Joshua and find that there are some surprising differences in implementation of even very fundamental features in each decoder.

@@

Additionally, the evaluation at stage (1) can return an estimate of the future cost to guide search. This is a technique used by the phrase-based decoder but not by the SCFG decoder.

Some feature functions do not override any of the stages. These feature is so intertwined with the search mechanism that their scores are set by search mechanism themselves. The feature function exist as an identifier to locate the score. Features functions such as the translation models, unknown word penalty,  generation models fall into this category. The input feature, which scores lattices and confusion network input is also hardcoded into the decoder.

@@

As with cdec and Moses' SCFG decoder, Joshua's decoding algorithm is an implementation of CKY+ \cite{chappelier1998generalized}.

Joshua makes only a single CKY pass over the source language sentence; it does not build a
``stateless'' hypergraph (or --LM forest, in the language of \citet{chiang2007hierarchical}). To
enable this, Joshua interleaves the extension of partially-constructed rules with the application of
complete rules while building the hypergraph, Figure~\ref{algorithm:joshua}:

@@

% The efficiency gained from this interleaving allows Joshua to decode with SCFGs with rich
% nonterminal sets \cite{galley2004whats,zollmann2006syntax}, although the time and memory
% requirements increase dramatically. Incorporating feature scores into the dot chart for more
% efficient decoding with SCFGs with rich nonterminal sets is a direction of interest.

\subsection{cdec}
In contrast to Moses and Joshua which use a one-pass search strategy, cdec uses a multi-pass search strategy. %That is, for each input sentence, all feature functions are evaluated when a hypothesis is created in Moses.
Each pass creates a hypergraph that is used as the input for the next pass. cdec evaluates stateless feature functions in the first pass, followed by stateful feature functions in the second.

@@

\subsection{Moses}
Moses implements supports for both dense and sparse features. Each feature must be identifiable as one or the other. The intuition behind this is that most systems will still contain dense features. Therefore, it would be more efficient to use existing dedicated data-structures to efficiently handle dense vectors. A sparse feature map was added to handle sparse scores.

@@

The results from the three decoders are comparable, given the same formalism and model data, and adjusting for scaling differences and constants. Figure~\ref{fig:model} shows the comparison of model scores and BLEU scores for a hierarchical model. Moses and cdec are roughly comparable in terms of speed. Also, Moses and cdec converges to roughly the same model scores, while Joshua makes more search errors, leading to overall lower BLEU scores.

@@

Figure~\ref{fig:glue} shows the improvement in cdec by changing the glue rule formulation from (2) to (1). The glue grammar for each decoder had varied since their inceptions. The convergence on using formulation (1) is the direct result of the comparison between the three decoders. (Result for cdec and Joshua in Figure~\ref{fig:model} were done with the improved glue grammar formulation.)

@@

@@

Statistical machine translation is a dynamic field which is constantly evolving. A major challenge for the Moses project, in common with many long-term software projects, is to maintain and update the the underlying framework and infrastructure to enable this evolution to continue. This paper documents the major changes that were made to Moses to enable this to happen and compare it two other decoders which implements the same decoding formalism.

%Even though each decoder was created independently, their output are comparable. We see that by doing an indepth analysis of the working of each decoder and comparing their model scores, we are able to transfer the best optmization techniques from each decoder to the others.
